{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2024-03-18 paper/main.tex
adding missing section

\noindent
{\bf Description of the evolutionary dynamics.}
Herein, we assume population members update their strategies based on social learning. 
To model these strategy updates, we consider a pairwise comparison process~\citep{traulsen2007pairwise}. 
This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to revise its strategy.
We refer to this player as the ``learner.'' 
With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (all reactive strategies have the same probability to be chosen). 
With the converse probability $1\!-\!\mu$, the learner randomly picks a ``role model'' from the population. 
The learner then compares its own payoff $\pi_\learner$ from the repeated game to the role model's payoff $\pi_\rolemodel$. 
The learner adopts the role model's strategy with a probability \(\varphi\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
\begin{equation} \label{Eq:rho}
    \varphi\left(\pi_\learner, \pi_\rolemodel\right) = \frac{1}{1\!+\! e^{- \!\beta\left(\pi_\rolemodel- \pi_\learner\right)}}.
\end{equation}
The selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability $1/2$. As the selection strength~$\beta$ increases, players are increasingly biased to imitate the role model only if it has the higher payoff. 
%In either case, we assume in the following that the role model's strategy can be identified.
%That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.

@@

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This finding is reminiscent of our earlier static results, which also suggested that changes in the game's benefits may have a negligible role on the stochastic stability of full cooperation.  
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection \(\beta \!<\! 1\). 
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

>>>>

\noindent
{\bf Description of the evolutionary dynamics.}
Herein, we assume population members update their strategies based on social learning. 
To model these strategy updates, we consider a pairwise comparison process~\citep{traulsen2007pairwise}. 
This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to revise its strategy.
We refer to this player as the ``learner''.
With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (all reactive strategies have the same probability to be chosen). 
With the converse probability $1\!-\!\mu$, the learner randomly picks a ``role model'' from the population. 
The learner then compares its own payoff $\pi_\learner$ from the repeated game to the role model's payoff $\pi_\rolemodel$. 
The learner adopts the role model's strategy with a probability \(\varphi\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
\begin{equation} \label{Eq:rho}
    \varphi\left(\pi_\learner, \pi_\rolemodel\right) = \frac{1}{1\!+\! e^{- \!\beta\left(\pi_\rolemodel- \pi_\learner\right)}}.
\end{equation}
The selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability $1/2$. As the selection strength~$\beta$ increases, players are increasingly biased to imitate the role model only if it has the higher payoff. 
%In either case, we assume in the following that the role model's strategy can be identified.
%That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.

@@

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This finding is reminiscent of our earlier static results, which also suggested that changes in the game's benefits may have a negligible role on the stochastic stability of full cooperation.  
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\! 1\)). This is not a coincidence: it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\
